Manuscript draft

### 2019-08-23 manuscript/manuscript draft.Rmd
updated manuscript draft

*Closure of fluxes* Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig. 3). Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

@@

*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 2).

```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("effect_of_latitude.png")
```

@@

*How does productivity relate to MAT and MAP?* MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature. We found no support for a saturation point of productivity with temperature.

MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux. For the majority of fluxes productivity is best predicted by a polynomial model. Productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease. The notable exception to this is GPP: the model indicates that GPP continues to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33. Data above this point is not available, but the model trend indicates that the saturation point for this model is around 5000mm mean annual precipitation.

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~. There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.

@@

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET, however all other fluxes showed a polynomial relationship with PET. We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm. There is evidence that productivity begins to decrease at values above 1800mm PET.

Solar radiation
Vapour pressure deficit?

@@

*What is the role of seasonality in explaining productivity?* Temperature seasonality is a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases. ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality. Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it. Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

@@

We find a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season. Length of growing season is a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it is not identified as a stronger predictor than MAT for any of the fluxes analysed.

>>>>

*Closure of fluxes* Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig. 2). Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.

@@

*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 3).

```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("effect_of_lat_transparent.png")
```

@@

*How does productivity relate to MAT and MAP?* MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig. 5). We found no support for a saturation point of productivity with temperature.

MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux. For the majority of fluxes productivity is best predicted by a polynomial model. Productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease (fig. 5). The notable exception to this is GPP: the model indicates that GPP continues to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33. Data above this point is not available, but the model trend indicates that the saturation point for this model is around 5000mm mean annual precipitation.

There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~ (fig.4). There was a significant additive effect for R~auto~. NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.

@@

We found a significant relationship between productivity and potential evapotranspiration for all fluxes. ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET, however all other fluxes showed a polynomial relationship with PET (fig. 5). We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm. There is evidence that productivity begins to decrease at values above 1800mm PET.

Vapour pressure deficit was a significant predictor of productivity for all fluxes. BNPP~fine root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig. 5). Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa. At values above 0.8 kPa, productivity began to decrease.

All fluxes, with the exception of R~root~, show a positive linear relationship with solar radiation. Solar radiation explains a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).

@@

```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.", fig.pos='H'}
  knitr::include_graphics("combined_plots.png")
```

*What is the role of seasonality in explaining productivity?* Temperature seasonality is a significant predictor of productivity. We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~. Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases (fig. 5). ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality. Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it. Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.

@@

We find a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig. 5). Length of growing season is a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it is not identified as a stronger predictor than MAT for any of the fluxes analysed.
